```rmd
---
title: "Oryza sativa IRGSP-1.0 Release 62 Annotation Package"
author: "Kabilan S"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{Oryza sativa IRGSP-1.0 Release 62 Annotation Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This package delivers Ensembl‑based gene, transcript, and exon annotations for 
*Oryza sativa* (rice) IRGSP‑1.0, Release 62. Gene IDs follow the **RAP‑DB** 
nomenclature (`OsXXgXXXXXXX`). MSU v7.0 functional descriptions are stored in 
the `gene_name` column, and NCBI Entrez Gene IDs in `entrezid`.

# Installation

From the source tarball:

```{r installing, eval = FALSE}
install.packages("EnsDb.Osativa.IRGSP.v62", repos = NULL, type = "source")
```

# Loading the package

```{r loading}
library(EnsDb.Osativa.IRGSP.v62)
edb <- tryCatch(
  EnsDb.Osativa.IRGSP.v62,
  error = function(e) {
    message("Database not available (maybe offline). Skipping example.")
    return(NULL)
  }
)
```

# Available columns and filters

```{r overview, eval = !is.null(edb)}
columns(edb)
supportedFilters(edb)
```

# Retrieve all genes

`genes()` returns a `GRanges` object with gene‑level information.

```{r genes, eval = !is.null(edb)}
gene_gr <- genes(edb)
head(gene_gr)
```

# Filter genes by ID

```{r filter_gene, eval = !is.null(edb)}
genes(edb, filter = GeneIdFilter("Os01g0100100"))
```

# Filter genes by biotype

```{r filter_gene_biotype, eval = !is.null(edb)}
genes(edb, filter = GeneBiotypeFilter("protein_coding"))
```

# Working with chromosome names

The database stores chromosome names as bare numbers (`1`, `2`, …). 
To add the UCSC‑style `chr` prefix:

```{r chromosomes, eval = !is.null(edb)}
gene_gr <- genes(edb)
seqlevelsStyle(gene_gr) <- "UCSC"
seqlevels(gene_gr)
```

Restrict to nuclear chromosomes:

```{r nuclear_chromosomes, eval = !is.null(edb)}
library(GenomeInfoDb)
gene_gr_nuc <- keepStandardChromosomes(gene_gr, pruning.mode = "coarse")
seqlevels(gene_gr_nuc)
```

# Session information

```{r session}
sessionInfo()
```

```
